### 2022-07-17 Shiny/Test/Flexdashboard test.Rmd
Plot and explanation. ifelse() rewrite.

---
title: "Charity analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl, flexdashboard, MetBrewer)

# Function to read html
read_proff6 <- function(file, start, end){
    read_html(file, encoding = "UTF-8") |> 
      html_node("#inner-frame") |> 
      html_table() |> 
      select( # removing empty column
        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
      mutate( # removing whitespace
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |> 
      filter( # removing duplicate table - no idea why it's necessary
        !row_number() > 176
      ) |>
      filter( # removing duplicate rows
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |> 
      pivot_longer( # tidying data
        all_of(end):all_of(start), names_to = "year") |> # using all_of()
      mutate( # changing to real NAs and turning years into numbers
        value = na_if(value, "-"),
        year = as.integer(year)) |> 
      mutate( # adding currency column
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |> 
      fill( # filling the currency column
        valutakode, .direction = "updown"
      ) |> 
      filter( # removing dates and redundant currency
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode") |> 
      mutate( # removing whitespace in numbers
        value = str_replace_all(value, "\\s", "")
      ) |> 
      mutate( # turning into numbers
        value = as.numeric(value)
      ) |> 
      mutate( # removing years as values
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |> 
      mutate( # values no longer in 1000
        value = value * 1000
      ) |> 
      distinct() |> # removing any remaining duplicate rows
      pivot_wider( # pivoting
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
      clean_names() |> # tidying names
      select(-lederlonn_i_hele_1000,
             -resultatregnskap_i_hele_1000,
             -balanseregnskap_i_hele_1000) |> # removing headings
      rename( # making it clear what lonn refers to
        lederlonn = lonn
      ) |> 
      arrange(desc(year)) # arranging by year
  }

# Function to read Excel file
widen_excel <- function(file){
    read_excel(file) |> 
      mutate(
        Value = Value * 1000
      ) |> 
      unite(
        entry, c(Category, Name), sep = "_"
      ) |> 
      select(-`Sub-category`) |> 
      pivot_wider(names_from = entry, values_from = Value) |> 
      clean_names()
}

# Shiny actions
observeEvent(input$click, { # wait for button click

    # creating data frame from html file
    htmldata <- reactive({
      read_proff6(input$html$datapath,
                  as.character(input$start), as.character(input$end))
      })

    # Creating data frame from Excel file
    exceldata <- reactive({
      widen_excel(input$excel$datapath)
    })

    # creating plots

    # ggplot as reactive function
    lonnplot <- reactive({
      ggplot(data = htmldata(), aes(year, lederlonn)) +
        geom_col() +
        scale_y_continuous(
          labels = scales::label_dollar(prefix = ""))
    })

    # Creating plotly plot
    output$plotlyplot <- renderPlotly(
      ggplotly(lonnplot())
    )

    # Creating ggplot
    output$htmlplot <- renderPlot(
      lonnplot()
    )

    # Creating fancier plot
    fancyplot <- reactive(
      htmldata() |> 
        select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
                    label_dollar(prefix = "? "))
      ) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
       subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
                         "",
                         "Note: not all values are denominated in the same currency"))
    )

    output$fancyplot <- renderPlotly(
      ggplotly(fancyplot())
    )
  })

```

Column {.sidebar}
-----------------------------------------------------------------------

### File upload

Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.

Then upload an Excel file in the set format.

Click 'Analyse' to analyse the data.

```{r}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")

```

Column 
-----------------------------------------------------------------------

### Plotly

```{r}
plotlyOutput("plotlyplot")
```

### ggplot
```{r}
plotOutput("htmlplot")
```

### Transparency
```{r}
plotlyOutput("fancyplot")
```