### 2022-07-17 Shiny/Test/Test data.Rmd
Adding Value Box.

---
title: "Test data"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-16'
output: 
  html_document: 
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Testing ideas for the Shiny app/dashboard.

# Background

## Libraries
```{r libraries}
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny,tidyverse, scales, rvest, janitor, plotly, readxl, MetBrewer)
```

## Functions

### HTML
```{r}
read_proff6 <- function(file, start, end){
    read_html(file, encoding = "UTF-8") |> 
      html_node("#inner-frame") |> 
      html_table() |> 
      select( # removing empty column
        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
      mutate( # removing whitespace
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |> 
      filter( # removing duplicate table - no idea why it's necessary
        !row_number() > 176
      ) |>
      filter( # removing duplicate rows
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |> 
      pivot_longer( # tidying data
        all_of(end):all_of(start), names_to = "year") |> # using all_of()
      mutate( # changing to real NAs and turning years into numbers
        value = na_if(value, "-"),
        year = as.integer(year)) |> 
      mutate( # adding currency column
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |> 
      fill( # filling the currency column
        valutakode, .direction = "updown"
      ) |> 
      filter( # removing dates and redundant currency
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode") |> 
      mutate( # removing whitespace in numbers
        value = str_replace_all(value, "\\s", "")
      ) |> 
      mutate( # turning into numbers
        value = as.numeric(value)
      ) |> 
      mutate( # removing years as values
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |> 
      mutate( # values no longer in 1000
        value = value * 1000
      ) |> 
      distinct() |> # removing any remaining duplicate rows
      pivot_wider( # pivoting
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
      clean_names() |> # tidying names
      select(-lederlonn_i_hele_1000,
             -resultatregnskap_i_hele_1000,
             -balanseregnskap_i_hele_1000) |> # removing headings
      rename( # making it clear what lonn refers to
        lederlonn = lonn
      ) |> 
      arrange(desc(year)) # arranging by year
}
```

### Excel wide
```{r}
widen_excel <- function(file){
  read_excel(file) |> 
    mutate(
      Value = Value * 1000
    ) |> 
  unite(
    entry, c(Category, Name), sep = "_"
  ) |> 
  select(-`Sub-category`) |> 
  pivot_wider(names_from = entry, values_from = Value) |> 
  clean_names()
}
```

## Loading data
```{r}
kreft_html <- read_proff6("Data/Kreftforeningen.html", "2002", "2021")
kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
```

## Viewing data

### `View()`
```{r view, eval=FALSE}
View(kreft_html)
View(kreft_excel)
```

### kabled
```{r}
kreft_html
kreft_excel
```

# Testing

## Data manipulation
```{r}
kreft_excel |> 
  filter(year == max(year))

kreft_html |> 
  filter(year == max(kreft_excel$year))

kreft_excel |> 
  left_join(kreft_html, by = "year")
```

## Visualisation ideas

### CEO compensation and currency testing
```{r}
compensation_p <- kreft_html |> 
  group_by(year) |> 
  mutate(
    leder_total = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE)
  ) |> 
  ggplot(aes(year, leder_total, fill = leder_total)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
ggplotly(compensation_p)

compensation_p2 <- kreft_html |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix =
                                    str_c(kreft_html$valutakode[1], " "))) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
ggplotly(compensation_p2)

# if else structure to check for uniform currency code?
length(unique(kreft_html$valutakode)) == 1

# Testing with uniform currency
if (length(unique(kreft_html$valutakode)) == 1) {
  kreft_html |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix =
                                    str_c(kreft_html$valutakode[1], " "))) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
} else {
    kreft_html |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix = "? ")) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
       subtitle = "Note: not all values are denominated in the same currency")
}

# Creating data frame with multiple currencies
currency_test <- kreft_html |> 
  mutate(
    valutakode = ifelse(year == 2019, "USD", valutakode)
  )

# Testing new data frame
if (length(unique(currency_test$valutakode)) == 1) {
  currency_test |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix =
                                    str_c(kreft_html$valutakode[1], " "))) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
} else {
    currency_test |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix = "? ")) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
       subtitle = "Note: not all values are denominated in the same currency")
}

# Could probably be done with two internal ifelse() functions

# ifelse() version

# Testing
currency_test |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = ifelse(length(unique(currency_test$valutakode)) == 1,
                    label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
                    label_dollar(prefix = "? "))
      ) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
       subtitle = ifelse(length(unique(currency_test$valutakode)) == 1,
                         element_blank(),
                         "Note: not all values are denominated in the same currency"))

# Original
kreft_html |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = ifelse(length(unique(kreft_html$valutakode)) == 1,
                    label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
                    label_dollar(prefix = "? "))
      ) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
       subtitle = ifelse(length(unique(kreft_html$valutakode)) == 1,
                         "",
                         "Note: not all values are denominated in the same currency"))
```

### Result
```{r}
arsresultat_p <- kreft_html |> 
  mutate(
    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
  ) |> 
  ggplot(aes(year, arsresultat, fill = profit)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = ""))
ggplotly(arsresultat_p)

arsresultat_p_met <- kreft_html |> 
  mutate(
    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
  ) |> 
  ggplot(aes(year, arsresultat, fill = profit)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  scale_fill_met_d("Tsimshian")
ggplotly(arsresultat_p_met)

ordinaert_p <- kreft_html |> 
  mutate(
    profit = ifelse(ordinaert_resultat >= 0, TRUE, FALSE)
  ) |> 
  ggplot(aes(year, ordinaert_resultat, fill = profit)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  theme(legend.position = "none")
ggplotly(ordinaert_p)

ordinaert_p_met <- kreft_html |> 
  mutate(
    profit = ifelse(ordinaert_resultat >= 0, TRUE, FALSE)
  ) |> 
  ggplot(aes(year, ordinaert_resultat, fill = profit)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  theme(legend.position = "none") +
  scale_fill_met_d("Tsimshian")
ggplotly(ordinaert_p_met)
```

## KPI ideas

### Fundraising ROI
```{r}
kreft_excel |> 
  mutate(
    KPI_ROI = expenses_fundraising / (revenue_private_contributions + revenue_membership_fees)
  ) |> 
  select(year, KPI_ROI)
```

### Administrative expenses
```{r}
kreft_excel |> 
  mutate(
    KPI_admin =
      str_c(
        format(
        (expenses_administrative / expenses_program_services *100), digits = 3
        ), "%"
        )
  ) |> 
  select(year, KPI_admin)

kreft_excel |> 
  mutate(
    KPI_admin =
      str_c(
        format(
        (expenses_administrative / expenses_program_services *100), digits = 3
        ), "%"
        )
  ) |> 
  filter(year == max(year)) |> 
  select(KPI_admin) |> 
  str_flatten()

kreft_excel |> mutate(
            KPI_admin =
              expenses_administrative / expenses_program_services *100) |> 
           filter(year == max(year)) |> 
          select(KPI_admin)
```

### Overhead
```{r}

```