### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Excel data quality tests and warnings.

---
title: "NGO analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using fontawesome for more up-to-date (and currently supported) fontawesome
# icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

Sidebar {.sidebar}
=======================================================================

### File upload

Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.

Then upload an Excel file in the set format.

Click 'Analyse' to analyse the data.

```{r sidebar}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")

```

Dashboard
=======================================================================

Row 
-----------------------------------------------------------------------

### KPI Fundraising
```{r}
valueBoxOutput("KPI_box_fundraising")
```

### KPI OPX
```{r}
valueBoxOutput("KPI_box_OPX")
```

### KPI Private growth
```{r}
valueBoxOutput("KPI_box_private")
```

### KPI Public growth
```{r}
valueBoxOutput("KPI_box_public")
```

### KPI Private/public proportion
```{r}
valueBoxOutput("KPI_box_proportion")
```

Row
-----------------------------------------------------------------------

### Annual result
```{r}
plotlyOutput("arsresultat_plotly")
```

### Balance
```{r}
plotlyOutput("balance_plotly")
```

Row 
-----------------------------------------------------------------------

### Expenses
```{r}
plotlyOutput("expenses_plotly")
```

### Revenues
```{r}
plotlyOutput("revenue_plotly")
```

Row 
-----------------------------------------------------------------------

### Transparency: CEO compensation
```{r}
plotlyOutput("CEO_plotly")
```

Details
=======================================================================

Row 
-----------------------------------------------------------------------

### Key Performance Indicators

The KPIs are based on the following calculations:

- **admin** is based on dividing - expenses_administrative / expenses_program_services *100
- **Fundraising ROI** - (revenue_private_contributions +
                 revenue_membership_fees - expenses_fundraising) /
              expenses_fundraising * 100
```{r}

```

Row 
-----------------------------------------------------------------------

### Plots
```{r}

```

```{r shiny, include=FALSE}
# Shiny actions
observe({ # wait for button click

  #---- Data frames ------------------------------------------------------------      
  # creating data frame from html file
  htmldata <- reactive({
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    widen_excel(input$excel$datapath)
  })

  #---- Value Boxes ------------------------------------------------------------

  # Creating Fundraising ROI value
  KPI_fundraising_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (revenue_private_contributions +
              revenue_membership_fees - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    }
  })

  # Creating OPX KPI
  KPI_OPX_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX = (revenue_private_contributions + revenue_investments +
          assets_cash_in_bank_and_cash_equivalents) /
          (liability_short_term_grants_payable +
            liability_revocable_endowments +
            liability_long_term_grants_payable) * 100
      ) |>
      # Just the single number
      pull() |>
      first()
    }
  })

  # Creating private donations growth
  KPI_private_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
    exceldata() |>
      # Calculating the two most recent years as a %
      transmute(
        growth = revenue_private_contributions[1] / revenue_private_contributions[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first()
    }
  })

  # Creating public grants growth
  KPI_public_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
    exceldata() |>
      # Calculating the two most recent years as a %
      transmute(
        growth = revenue_public_grants[1] / revenue_public_grants[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first()
    }
  })

  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        transmute(
          proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
        ) |>
        pull() |>
        first()
    }
  })

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box_fundraising <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_fundraising_num()),
      caption = "Fundraising ROI",
      # Colour depends on value
      color = if (is.na(KPI_fundraising_num())) {
        "primary"
      } else if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-donate"
    )
  })

  # KPI_OPX
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-seedling"
    )
  })

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 100) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-hands-helping"
    )
  })

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 100) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-hand-holding-usd"
    )
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      color = "info",
      icon = "fas fa-percentage"
    )
  })

  #---- Plots --------------------------------------------------------------------    
  #----CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    # Testing for the presence of data
    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      # Creating plot if any data exist
      # Note that the plot will still look empty, but have a legend and scales,
      # if all values are explicit zeroes
      htmldata() |>
        # preparing for pivoting
        select(
          year, lederlonn, leder_annen_godtgjorelse
        ) |>
        # Compensation in ordinary numbers
        mutate(
          lederlonn = lederlonn * 1000,
          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
        ) |>
        # renaming for the purposes of the plot
        rename(
          Year = year,
          Salary = lederlonn,
          `Other compensation` = leder_annen_godtgjorelse
        ) |>
        # pivoting to join the two values to use fill
        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
        ggplot(aes(Year, value, fill = Compensation)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # pretty colours
        scale_fill_met_d("Hokusai2", direction = -1) +
        labs(
          x = "", y = "Total compensation", title = "CEO compensation"
        )
    }
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Annual result---------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        ggplot(aes(year, arsresultat, fill = annual_profit)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Annual result (in thousands)"
        )
    }
  })

  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Balance--------------------------------------------------------------------    
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Balance (in thousands)",
        )
    }
  })

  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Expenses---------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum())
      ) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot())
  )

  #----Revenue------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum())
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

  output$revenue_plotly <- renderPlotly(
    ggplotly(revenue_plot())
  )

  #----Reactive functions-------------------------------

  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })

  # Reactive function to check for currency and use currency code or ?
  currency_scale <- reactive({
    scale_y_continuous(
      # Checking for uniform currency code
      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
        # Using the currency code if uniform
        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
        # alternative prefix if multiple currencies in the data
        label_dollar(prefix = "? ")
      )
    )
  })

  # Reactive function to create currency caption
  currency_caption <- reactive({
    # Checking for uniform currency code
    if (length(unique(htmldata()$valutakode)) == 1) {
      # Empty caption if uniform
      ""
      # Warning if multiple currencies
    } else {
      "Note: not all values are denominated in the same currency"
    }
  })

  # Reactive function to create plotly caption
  # (ggplot captions and subtitles are lost in the conversion)
  # (shamelessly stolen from the Internet, with minor modifications)
  plotly_caption <- reactive({
    list(
      x = 1, y = -0.1, text = currency_caption(),
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 11)
    )
  })

  # Reactive function to create the maximum value for revenue and expenses plots
  maximum <- reactive({
    # reading the Excel sheet anew
    read_excel(input$excel$datapath) |>
      # removing the firm and sub-category columns
      select(-Firm, -`Sub-category`) |>
      # filtering to retain only revenue and expenses entries
      filter(Category == "Revenue" | Category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(Year, Category) |>
      summarise(
        sum = sum(Value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  })
  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)

```

```{r non-reactive functions, include=FALSE}
# Function to read html file
read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    arrange(desc(year)) # arranging by year
}

# Test version of function that creates multiple data problems
# read_proff <- function(file, start, end) {
#   read_html(file, encoding = "UTF-8") |>
#     html_node("#inner-frame") |>
#     html_table() |>
#     select( # removing empty column
#       REGNSKAPSPERIODE:all_of(start)
#     ) |> # using all_of()
#     mutate( # removing whitespace
#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
#     ) |>
#     filter( # removing duplicate table - no idea why it's necessary
#       !row_number() > 176
#     ) |>
#     filter( # removing duplicate rows
#       !grepl("Lukk", REGNSKAPSPERIODE)
#     ) |>
#     pivot_longer( # tidying data
#       all_of(end):all_of(start),
#       names_to = "year"
#     ) |> # using all_of()
#     mutate( # changing to real NAs and turning years into numbers
#       value = na_if(value, "-"),
#       year = as.integer(year),
#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
#     ) |>
#     fill( # filling the currency column
#       valutakode,
#       .direction = "updown"
#     ) |>
#     filter( # removing dates and redundant currency
#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
#         REGNSKAPSPERIODE != "Valutakode"
#     ) |>
#     mutate( # removing whitespace in numbers
#       value = str_replace_all(value, "\\s", ""),
#       value = as.numeric(value),
#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
#     ) |>
#     distinct() |> # removing any remaining duplicate rows
#     pivot_wider( # pivoting
#       names_from = REGNSKAPSPERIODE, values_from = value
#     ) |>
#     clean_names() |> # tidying names
#     select(
#       -lederlonn_i_hele_1000,
#       -resultatregnskap_i_hele_1000,
#       -balanseregnskap_i_hele_1000
#     ) |> # removing headings
#     rename( # making it clear what lonn refers to
#       lederlonn = lonn
#     ) |>
#     # Creating new variables for plots
#     mutate(
#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
#     ) |>
#     # bad test - REMOVE
#     mutate(
#       valutakode = ifelse(year == 2019, "USD", valutakode),
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA
#     ) |>
#     arrange(desc(year)) # arranging by year
# }

# Function to read Excel file
widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}

# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}

# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
    scale_fill_met_d("Tsimshian",
      direction = ifelse(
        mean(variable,
          na.rm = TRUE
        ) == 1, -1, 1
      )
    )
  }

# Function to print a 'no information' plot, if necessary
no_information <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "No information available",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}

# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "Something is wrong\nwith the Excel data.",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```