### 2022-07-23 Shiny/Test/Colourful plots.Rmd
Code hiding as default.

---
title: "Colourful plot tests"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny,tidyverse, scales, rvest, janitor, plotly, readxl, MetBrewer)
```

```{r data}

read_proff8 <- function(file, start, end){
    read_html(file, encoding = "UTF-8") |> 
      html_node("#inner-frame") |> 
      html_table() |> 
    # removing empty column
      select( 
        REGNSKAPSPERIODE:all_of(start)) |>
    # removing whitespace
      mutate( 
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |> 
    # removing duplicate table - no idea why it's necessary
      filter(
        !row_number() > 176
      ) |>
    # removing duplicate rows
      filter( 
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |> 
    # tidying data
      pivot_longer( 
        all_of(end):all_of(start), names_to = "year"
        ) |> 
    # changing to real NAs and turning years into numbers
    # picking out the currency
      mutate( 
        value = na_if(value, "-"),
        year = as.integer(year), 
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |> 
    # filling the currency column
      fill( 
        valutakode, .direction = "updown"
      ) |> 
    # removing dates and redundant currency
      filter( 
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode") |> 
    # removing white space in numbers
    # removing years as numbers
      mutate( 
        value = str_replace_all(value, "\\s", ""),
        value = as.numeric(value),
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |> 
    # removing any remaining duplicate rows
      distinct() |> 
    # pivoting
      pivot_wider( 
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
    # tidying names
      clean_names() |> 
    # removing headings
      select(-lederlonn_i_hele_1000,
             -resultatregnskap_i_hele_1000,
             -balanseregnskap_i_hele_1000) |> 
    # making it clear what lonn refers to
      rename( 
        lederlonn = lonn
      ) |> 
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |> 
    # arranging by year
      arrange(desc(year)) 
}

widen_excel <- function(file){
  read_excel(file) |> 
  unite(
    entry, c(Category, Name), sep = "_"
  ) |> 
  select(-`Sub-category`) |> 
  pivot_wider(names_from = entry, values_from = Value) |> 
  clean_names()
}

kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")

kreft_html <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")

```

## Four colours
```{r}
palettes <- colorblind_palettes

for (i in 1:length(palettes)) {
  print(kreft_excel |> 
    transmute(
      Year = as.factor(year),
      Other = expenses_other,
      Administrative = expenses_administrative,
      Fundraising = expenses_fundraising,
      `Program services` = expenses_program_services
    ) |> 
    pivot_longer(c(Other:`Program services`), names_to = "Expenses") |> 
    mutate(
      Expenses = ordered(Expenses, levels =
                           c("Program services", "Administrative",
                             "Fundraising", "Other")) 
    ) |> 
    ggplot(aes(Year, value, fill = Expenses)) +
    geom_col() +
    coord_flip() +
    scale_y_continuous(labels = label_dollar()) +
    scale_fill_met_d(palettes[i]) +
    labs(x = "Year", y = "", title = palettes[i]))
}
```

## Two colours
```{r}
for (i in 1:length(palettes)) {
  print(kreft_html |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(
    labels = scales::label_dollar(prefix =
                                    str_c(kreft_html$valutakode[1], " "))) +
  scale_fill_met_d(palettes[i]) +
  labs(x = "Year", y = "Total compensation", title = palettes[i]))
}
```