### 2022-07-23 Report/KPIs and Visualisations.Rmd
Creating Report folder. Uploading a document about KPIs and visualisations (with code). Adding info about the KPIs and plots in the dashboard.

---
title: "KPIs and Visualisations"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-23'
output: 
  html_document: 
    toc: yes
    toc_depth: 5
    code_folding: show
---

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
```

>Sannsynligvis mest oversiktlig hvis man kjører "Knit".

Laster opp dette dokumentet som informasjon om KPIer og visualiseringer -- samt muligens som inspirasjon for visualiseringer og utregninger i rapporten.

Legg merke til at innenfor Shiny så er de fleste objekter ikke objekter men funksjoner, dvs. at objektet `htmldata` her heter `htmldata()` i Shiny-koden. I tillegg er mesteparten av elementene kodet som "reaktive funksjoner", dvs. pakket inn i `reactive({})`. Den delen av koden kan fjernes i et vanlig R Markdown-dokument. De forskjellige datatestene er nok heller ikke nødvendige.

En idé om hvordan pakken [MetBrewer](https://github.com/BlakeRMills/MetBrewer), som jeg har brukt for farger, virker i praksis kan man se hvis man kjører "Knit" på dokumentet "Colourful Plots" i Shiny/Test.

Legg merke til at ingen av koden, unntatt setup-chunken kjøres når man knitter, men koden vises. (`knitr::opts_chunk$set(echo = TRUE, eval = FALSE)`)

*Si ifra hvis du/dere har spørsmål!*

## KPIs

### Key Performance Indicators

>Her har jeg brukt enkle LaTeX mattefunsksjoner for å vise utregningene.

The KPIs are based on the following calculations, using data derived from the Excel sheet:

- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
  - A basic measure of the ROI of direct fundraising expenses.
- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
  - A measure of the relationship between the organisation's liquid assets and its obligations.
- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
  - A calculation of the percentage change in private contributions based on the two most recent years.
- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
  - A calculation of the percentage change in public grants received based on the two most recent years.
- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.

### Code

#### Fundraising ROI
```{r}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (revenue_private_contributions +
            revenue_membership_fees - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

#### OPX KPI
```{r}
# Creating OPX KPI
KPI_OPX_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX = (revenue_private_contributions + revenue_investments +
          assets_cash_in_bank_and_cash_equivalents) /
          (liability_short_term_grants_payable +
            liability_revocable_endowments +
            liability_long_term_grants_payable) * 100
      ) |>
      # Just the single number
      pull() |>
      first()
  }
})
```

#### private donations growth
```{r}
# Creating private donations growth
KPI_private_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_private_contributions[1] - revenue_private_contributions[2]) /
            revenue_private_contributions[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first()
  }
})
```

#### public grants growth
```{r}
# Creating public grants growth
KPI_public_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_public_grants[1] - revenue_public_grants[2]) /
            revenue_public_grants[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first()
  }
})
```

#### public/private proportion
```{r}
# Creating public/private proportion
KPI_proportion <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      transmute(
        proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
      ) |>
      pull() |>
      first()
  }
})
```

## Plots

The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

### Code

#### Årsresultat
```{r}
arsresultat_plot <- reactive({
  # testing for the presence of data
  if (sum(!is.na(htmldata()$arsresultat)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    htmldata() |>
      ggplot(aes(year, arsresultat, fill = annual_profit)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale() +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata()$annual_profit) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Annual result (in thousands)"
      )
  }
})

# Creating plotly output
output$arsresultat_plotly <- renderPlotly(
  ggplotly(arsresultat_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)
```

#### Balance
```{r}
#----Balance--------------------------------------------------------------------
balance_plot <- reactive({
  # testing for the presence of data
  if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    htmldata() |>
      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale() +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata()$balance_positive) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Balance (in thousands)",
      )
  }
})

# Creating plotly output
output$balance_plotly <- renderPlotly(
  ggplotly(balance_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)
```

#### Expenses
```{r}
expenses_plot <- reactive({
  # testing for data quality
  if (datatest() == TRUE) {
    # printing 'bad data' plot if data are bad
    bad_data()
  } else {
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum())
      ) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)")
  }
})

# Creating plotly output
output$expenses_plotly <- renderPlotly(
  ggplotly(expenses_plot())
)
```

#### Revenue
```{r}
revenue_plot <- reactive({
  # testing for data quality
  if (datatest() == TRUE) {
    # printing 'bad data' plot if data are bad
    bad_data()
  } else {
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum())
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)")
  }
})

# Creating plotly output
output$revenue_plotly <- renderPlotly(
  ggplotly(revenue_plot())
)
```

#### CEO compensation
```{r}
#----CEO compensation---------------------------------------------------------
CEO_plot <- reactive({
  # Testing for the presence of data
  if (sum(!is.na(htmldata()$lederlonn)) == 0 &
    sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    # Creating plot if any data exist
    # Note that the plot will still look empty, but have a legend and scales,
    # if all values are explicit zeroes
    htmldata() |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale() +
      # pretty colours
      scale_fill_met_d("Hokusai2", direction = -1) +
      labs(
        x = "", y = "Total compensation", title = "CEO compensation"
      )
  }
})

# Creating plotly output
output$CEO_plotly <- renderPlotly(
  ggplotly(CEO_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)
```

## Background functions

### Code

#### Reactive functions

##### datatest
```{r}
# Reactive function to test for Excel data quality (basic)
datatest <- reactive({
  # let me know if anyone has a better idea for testing
  if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
    TRUE
  } else {
    FALSE
  }
})
```

##### currency_scale
```{r}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

##### plotly_caption
```{r}
# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

##### maximum
```{r}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  # reading the Excel sheet anew
  read_excel(input$excel$datapath) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
})
```

#### Static functions

##### Creating the necessary data
```{r data}
# Function to read html file
read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    arrange(desc(year)) # arranging by year
}

# Creating data frame
htmldata <- read_proff("Data/Kreftforeningen.html", "2002", "2021")

# Function to read Excel file
widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}

# Creating data frame
exeldata <- widen_excel("Data/OPXfinans_final.xlsx")
```

##### KPI_to_string
```{r}
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}
```

##### scale_fill_green
```{r}
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}
```

##### no_information
```{r}
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "No information available",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

##### bad_data
```{r}
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "Something is wrong\nwith the Excel data.",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```