### 2022-07-26 Old/Dag 3 heck.Rmd
Cleanup.

---
title: "Data analysis and data transformation"
author: "Henrik Rui"
date: '2022-06-29'
output:
  pdf_document: default
  html_document: default
---

#Tinytex
#Setter eval til FALSE for aa ikke evaluere kode når man lager dokumentet

```{r, eval=FALSE}
install.packages("tinytex")
library(tinytex)
install_tinytex()

```