### 2022-07-26 Old/Shiny/Test/Facet test.Rmd

---
title: "Facet test"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny,tidyverse, scales, rvest, janitor, plotly, readxl, MetBrewer, cowplot)
```

```{r}
revenue <- kreft_excel |> 
  # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year                    = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees`       = revenue_membership_fees,
        `Public grants`         = revenue_public_grants,
        Investments             = revenue_investments,
        `Operational activities`= revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other                   = revenue_other
      ) |> 
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other), names_to = "Revenue source"
      ) |> 
  ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
scale_y_continuous(labels = label_dollar(prefix = "NOK "),
                         limits = c(NA, maximum)) +
  scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)")
revenue
expenses <- kreft_excel |> 
  transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |> 
  ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(labels = label_dollar(prefix = "NOK "),
                         limits = c(NA, maximum)) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)")

expenses

maximum

?scale_y_continuous()

expenses

maximum <- read_excel("Data/OPXfinans_final.xlsx") |> 
  select(-Firm, -`Sub-category`) |> 
  filter(Category == "Revenue" | Category == "Expenses") |> 
  group_by(Year, Category) |> 
  summarise(
    sum = sum(Value, na.rm = TRUE)
  ) |> 
  pull() |> 
  max()

maximum

excel_long

bare
filtered <- bare |> 
  filter(Category == "Revenue" | Category == "Expenses")

filtered
filtered <- filtered |>
  select(-Firm, -`Sub-category`)

filtered |> 
  ggplot(aes(as.factor(Year), Value, fill = Name)) +
  geom_col() +
  facet_wrap(~Category) +
  scale_y_continuous(labels = label_dollar()) +
  scale_fill_manual(values = met.brewer("Hokusai3", 10))

maximum <- filtered |> 
  group_by(Year, Category) |> 
  summarise(
    sum = sum(Value, na.rm = TRUE)
  ) |> 
  pull() |> 
  max()

filtered |> 
  group_by(Year, Category) |> 
  summarise(
    sum = sum(Value, na.rm = TRUE)
  )

```

# Data
```{r}
widen_excel <- function(file){
  read_excel(file) |> 
  unite(
    entry, c(Category, Name), sep = "_"
  ) |> 
  select(-`Sub-category`) |> 
  pivot_wider(names_from = entry, values_from = Value) |> 
  clean_names()
}

kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
```