### 2022-07-26 Old/Shiny/Test/Reading OPX table.Rmd

---
title: "Reading OPX table"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
care <- read_html("Shiny/Test/Data/NGO Profile Page - OPX.webarchive") |> 
  html_element("body > opx-root > opx-ngo-profile-sidenav > opx-sidenav > div > mat-drawer-container > mat-drawer-content > div > opx-ngo-profile > opx-ngo-financial > div") |> 
  html_table()

care
care2 <- care |> 
  rename(
    Name = X1,
    `2020` = X2,
    `2019` = X3,
    `2018` = X4,
    `2017` = X5
  ) |> 
  pivot_longer(`2020`:`2017`, names_to = "Year")
care2 |> 
  mutate(
    value = (str_replace_all(value, "\\s", "")),
    value = as.numeric(value),
    Year = as.integer(Year)
  ) |> 
  arrange(desc(Year)) |> 
  write_excel_csv("Shiny/Test/Data/care.csv")
```