### 2022-07-26 Report/Draft Dashboard section.Rmd
Updating with info on report; updating some of the code as well.

---
title: "Dashboard section"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-23'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

>Ca. 8 sider.

# The Dashboard

One of the outputs we have created is an example dashboard that calculates some KPIs and creates a few visualisations based on user-uploaded data from Proff.no and the Excel template we developed. The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'under the hood' was created using Shiny. Additionally, the dashboard provides the user with the option of downloading a pdf version of the analysis, using a parameterised R Markdown document.

As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed:
```{r dashboard_packages, eval=FALSE}
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
```

The dashboard consists of three main parts: a sidebar, the dashboard proper, and a 'Details' tab that provides basic information about the KPIs and plots.

## The Sidebar

The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data. (Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)
```{r dashboard_sidebar, eval=FALSE}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")
```

Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded:
```{r dashboard_file_validation, eval=FALSE}
validate(need(tools::file_ext(input$html$datapath) == "html",
                  "Please upload an html file."))
validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
                  "Please upload an Excel file."))
```

If the validation is successful the files are read using the functions developed for this purpose.

## The dashboard proper

### Key Performance Indicators

The main section of the dashboard has a row-based layout, with a top row containing five 'value boxes' displaying KPIs derived from data in the Excel sheet, calculated and formatted in separate reactive functions. As Flexdashboard fails particularly gracelessly when errors occur with the value boxes, each function relies on some basic checks carried out by separate functions created for this purpose. In the case of both this is a simple function that checks if any of the numeric columns contain characters (in which case the data frame is probably hopelessly corrupted):
```{r dashboard_datatest, eval=FALSE}
datatest <- reactive({
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
```

If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
```{r dashboard_KPI_calculation, eval=FALSE}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

The value box then also relies on a function that converts the number into a percentage string, checking in the process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
```{r dashboard_KPI_string, eval=FALSE}
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
  # avoid passing an NA to subsequent operations.)
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}
```

In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour and icon of the box based on the calculated number. The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
```{r dashboard_valuebox, eval=FALSE}
# Creating value box for fundraising ROI
output$KPI_box_fundraising <- renderValueBox({
  valueBox(
    value = KPI_to_string(KPI_fundraising_num()),
    caption = "Fundraising ROI",
    # Colour depends on value
    color = if (is.na(KPI_fundraising_num())) {
      "primary"
    } else if (KPI_fundraising_num() < 100) {
      "danger"
    } else if (KPI_fundraising_num() < 200) {
      "warning"
    } else {
      "success"
    },
    # icon depends on whether or not the value is NA
    icon = if (is.na(KPI_fundraising_num())) {
      "fa-solid fa-circle-xmark"
    } else {
      "fas fa-donate"
    }
  )
})
```

All the boxes also include, as mentioned, a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1. All boxes use the theme colour 'primary' and the icon 'fa-solid fa-circle-xmark' if the value is NA.
```{r dashboard_fontawesome, eval=FALSE}
# Using fontawesome for more up-to-date (and currently supported) fontawesome
# icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

### Plots

Below the row of KPI boxes, the dashboard contains three more rows containing five plots, three based on Proff.no data and two based on Excel data. The html-based plots include a basic test of whether the necessary data are available, as some or all may sometimes be missing, printing a 'no information' plot using a separate function if the data is missing.
```{r html_datatest, eval=FALSE}
# testing for the presence of data
if (sum(!is.na(htmldata()$arsresultat)) == 0) {
  # Printing 'no information' plot if no data
  no_information()
} else {
  ggplot() # create plot
}
```

The 'no information' plot is an essentially empty plot displaying a notice to the user, and is created through a function that passes label information into a more general blank plot function:
```{r dashboard_no_information, eval=FALSE}
## no_information plot ---------------------------------------------------------
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  empty_plot("No information available")
}

## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

The top row contains two plots showing the annual result and balance data for the years covered by Proff.no. These are fairly simple bar plots, including a `fill` parameter colouring the bars depending on whether the numbers are positive or negative. They do, however, rely on two simple additional functions, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, or, otherwise, uses a question mark instead of a currency code, both using `scales::label_dollar()`.
```{r dashboard_currency, eval=FALSE}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

A second function is used for the `fill` parameter in order to (1) ensure that positive numbers are always green, irrespective of the presence of negative numbers, and vice versa, and (2) to use slightly nicer colours than the default ggplot2 ones, making use of the 'Tsimshian' palette from Blake R. Mills' creative [MetBrewer](https://github.com/BlakeRMills/MetBrewer) package.

The plots are converted to plotly using `ggplotly()`. As subtitles and captions do not survive the conversion, a separate function outputs a plotly annotation if the data contains multiple currency codes:
```{r dashboard_plotly, eval=FALSE}
# Creating plotly output
output$arsresultat_plotly <- renderPlotly(
  ggplotly(arsresultat_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)

# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

Below these plots are two plots based on Excel data. These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a function that passes label information to the `empty_plot()` function:
```{r dashboard_excel_datatest, eval=FALSE}
## bad_data plot ---------------------------------------------------------------
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  empty_plot("Something is wrong\nwith the Excel data.")
}
```

Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting - in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order. Excel data are presumed to be supplied in NOK.
```{r dashboard_excelplot, eval=FALSE}
exceldata() |>
  # Renaming for the purposes of the plot + dropping other variables
  transmute(
    Year = as.factor(year),
    `Private contributions` = revenue_private_contributions,
    `Membership fees` = revenue_membership_fees,
    `Public grants` = revenue_public_grants,
    Investments = revenue_investments,
    `Operational activities` = revenue_operational_activities,
    `Commercial activities` = revenue_commercial_activities,
    Other = revenue_other
  ) |>
  # Pivoting to use fill
  pivot_longer(
    c(`Private contributions`:Other),
    names_to = "Revenue source"
  ) |>
  # Ordering for the plot
  mutate(
    `Revenue source` =
      ordered(`Revenue source`,
        levels =
          c(
            "Other", "Commercial activities",
            "Operational activities", "Investments",
            "Membership fees", "Private contributions",
            "Public grants"
          )
      )
  ) |>
  ggplot(aes(Year, value, fill = `Revenue source`)) +
  geom_col() +
  scale_y_continuous(
    # Excel data are assumed to be in NOK; if this changes, code based on
    # the currency_code() function may be adapted
    labels = label_dollar(prefix = "NOK "),
    # Basing lower limit on the data (not suppressing erroneous negative
    # numbers), and the upper limit on the maximum value in revenues and
    # expenses, to match the two plots
    limits = c(NA, maximum())
  ) +
  # Specifying the number of values, as the base palette only includes 6
  scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
  labs(x = "", y = "", title = "Revenue (in thousands)")
```

The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data. The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals. (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
```{r dashboard_maximum, eval=FALSE}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  # reading the Excel sheet anew
  read_excel(input$excel$datapath) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
})
```

All the plots based on the Excel data make use of the "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories. The same function is used to generate the text colour in the 'no information' and 'bad data' plots.

Below these two plots is a plot showing the total reported CEO compensation in the years covered by the html data, where this information is available.

## Details tab

Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
```{r dashboard_title_creation, eval=FALSE}
# NGO title ==================================================================
NGO <- reactive({
  # validating the file again
  validate(need(
    tools::file_ext(input$excel$datapath) == "xlsx",
    "Please upload an Excel file."
  ))
  # fetching the name
  read_excel(input$excel$datapath) |>
    # Turning the needed column into a list
    pull(Firm) |>
    # discarding all NAs in case the name is somewhere
    purrr::discard(is.na) |>
    # choosing the first available name
    first()
})

title <- reactive({
  # testing if NGO() is NA
  if (is.na(NGO())) {
    # Printing warning if NA
    "Name of NGO missing"
    # Printing name of NGO
  } else {
    NGO()
  }
})
```

The title is output using `htmlOutput()` to use some basic html styling:
```{r dashboard_title, eval=FALSE}
# Creating NGO title using html formatting (colour based on MetBrewer
# "Hokusai2")
output$title <- renderText({
  paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
})
```

## The parameterised R Markdown report

The dashboard also allows the user to download a pdf version of the analysis through clicking on a button. A `downloadHandler()` function passes the necessary parameters to the R Markdown document and displays a temporary notification while the file is being created. The function includes temporary file creation arguments that should enable it to function on e.g. shinyapps.io and similar server contexts, but this has not been tested.
```{r}
## Download and parameters -----------------------------------------------------
output$report <- downloadHandler(

  # Based on code in Mastering Shiny
  filename = "report.pdf",
  content = function(file) {
    tempReport <- file.path(tempdir(), "report.Rmd")
    file.copy("report.Rmd", tempReport, overwrite = TRUE)

    # parameters
    params <- list(
      htmlfile = input$html$datapath,
      excelfile = input$excel$datapath,
      start = input$start,
      end = input$end
    )

    # notification
    id <- showNotification(
      "Generating report...",
      duration = NULL,
      closeButton = FALSE
    )
    on.exit(removeNotification(id), add = TRUE)

    # R Markdown render
    rmarkdown::render(tempReport,
      output_file = file,
      params = params,
      envir = new.env(parent = globalenv())
    )
  }
)
```

The code in the R Markdown document is very similar to that of the dashboard, but structured in a more conventional way. Data warnings are here text rather than blank plots, and the KPIs are integrated into the text instead of being presented in value boxes.