### 2022-07-28 Old/Shiny/Test/HTML tests.Rmd
HTML function testing.

---
title: "HTML tests"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-07-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Part 1
read_proff_read <- function(file) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")
}

# Part 2 - version with integrated test
read_proff_complete <- function(object, start, end) {
  if (as.character(all_of(start)) %in% names(object) &&
    as.character(all_of(end)) %in% names(object)) {
    object |>
      # removing empty column, using all_of(argument)
      select(
        REGNSKAPSPERIODE:all_of(start)
      ) |>
      # removing white space
      mutate(
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |>
      # removing duplicate table sometimes created by read_html()
      filter(
        !row_number() > 176
      ) |>
      # removing duplicate rows
      filter(
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |>
      # tidying data using pivot_longer() and all_of(argument)
      pivot_longer(
        all_of(end):all_of(start),
        names_to = "year"
      ) |>
      # creating a currency code column and turning current currency codes into
      # NAs
      mutate(
        # changing to real NAs
        value = na_if(value, "-"),
        # turning years into numbers
        year = as.integer(year),
        # creating a currency code column and turning current currency codes into
        # NAs
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |>
      # filling the currency column
      fill(
        valutakode,
        .direction = "updown"
      ) |>
      # removing dates and redundant currency
      filter(
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode"
      ) |>
      mutate(
        # removing white space in numbers
        value = str_replace_all(value, "\\s", ""),
        # turning numbers into true numbers
        value = as.numeric(value),
        # removing values that are years
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |>
      # removing any remaining duplicate rows
      distinct() |>
      # pivoting to create 'tidy' variables
      pivot_wider(
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |>
      # tidying names
      clean_names() |>
      # removing headings from the table
      select(
        -lederlonn_i_hele_1000,
        -resultatregnskap_i_hele_1000,
        -balanseregnskap_i_hele_1000
      ) |>
      # making it clear what lonn refers to
      rename(
        lederlonn = lonn
      ) |>
      # Creating new variables for plots
      mutate(
        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
      ) |>
      # arranging by year for practical display
      arrange(desc(year))
  } else {
    validate("Please ensure you have entered the correct start and end years")
  }
}

# creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))

    # Testing
    html_test <- read_proff_test(input$html$datapath)

    # Creating data frame

    read_proff_complete(
      html_test,
      as.character(input$start), as.character(input$end)
    )
  })
```

```{r}
html_testing <- function(file, start, end) {
  # Reading file and node
  htmltest <- read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")

  # Checking of node exists
  if (class(htmltest) == "xml_node") {
    htmltable <- htmltest |>
      html_table()
  } else {
    validate("table error")
  }

  # Basic check if the table is correct
  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
    htmlready <- htmltable
  } else {
    validate("regnskap error")
  }

  # Testing if the correct years have been entered
  if (as.character(all_of(start)) %in% names(htmlready) &&
    as.character(all_of(end)) %in% names(htmlready)) {
    # Creating the data frame
    read_proff(file, start, end)
  } else {
    validate("Please ensure you have entered the correct start and end years.")
  }
}

read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |> 
      # removing empty column, using all_of(argument)
      select(
        REGNSKAPSPERIODE:all_of(start)
      ) |>
      # removing white space
      mutate(
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |>
      # removing duplicate table sometimes created by read_html()
      filter(
        !row_number() > 176
      ) |>
      # removing duplicate rows
      filter(
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |>
      # tidying data using pivot_longer() and all_of(argument)
      pivot_longer(
        all_of(end):all_of(start),
        names_to = "year"
      ) |>
      # creating a currency code column and turning current currency codes into
      # NAs
      mutate(
        # changing to real NAs
        value = na_if(value, "-"),
        # turning years into numbers
        year = as.integer(year),
        # creating a currency code column and turning current currency codes into
        # NAs
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |>
      # filling the currency column
      fill(
        valutakode,
        .direction = "updown"
      ) |>
      # removing dates and redundant currency
      filter(
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode"
      ) |>
      mutate(
        # removing white space in numbers
        value = str_replace_all(value, "\\s", ""),
        # turning numbers into true numbers
        value = as.numeric(value),
        # removing values that are years
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |>
      # removing any remaining duplicate rows
      distinct() |>
      # pivoting to create 'tidy' variables
      pivot_wider(
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |>
      # tidying names
      clean_names() |>
      # removing headings from the table
      select(
        -lederlonn_i_hele_1000,
        -resultatregnskap_i_hele_1000,
        -balanseregnskap_i_hele_1000
      ) |>
      # making it clear what lonn refers to
      rename(
        lederlonn = lonn
      ) |>
      # Creating new variables for plots
      mutate(
        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
      ) |>
      # arranging by year for practical display
      arrange(desc(year))
}
```