### 2022-07-28 Old/Shiny/Test/HTML tests.Rmd
Found a problem with the html code. fml.

---
title: "HTML tests"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-07-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
```

```{r}
# Part 1
read_proff_read <- function(file) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")
}

# Part 2 - version with integrated test
read_proff_complete <- function(object, start, end) {
  if (as.character(all_of(start)) %in% names(object) &&
    as.character(all_of(end)) %in% names(object)) {
    object |>
      # removing empty column, using all_of(argument)
      select(
        REGNSKAPSPERIODE:all_of(start)
      ) |>
      # removing white space
      mutate(
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |>
      # removing duplicate table sometimes created by read_html()
      filter(
        !row_number() > 176
      ) |>
      # removing duplicate rows
      filter(
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |>
      # tidying data using pivot_longer() and all_of(argument)
      pivot_longer(
        all_of(end):all_of(start),
        names_to = "year"
      ) |>
      # creating a currency code column and turning current currency codes into
      # NAs
      mutate(
        # changing to real NAs
        value = na_if(value, "-"),
        # turning years into numbers
        year = as.integer(year),
        # creating a currency code column and turning current currency codes into
        # NAs
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |>
      # filling the currency column
      fill(
        valutakode,
        .direction = "updown"
      ) |>
      # removing dates and redundant currency
      filter(
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode"
      ) |>
      mutate(
        # removing white space in numbers
        value = str_replace_all(value, "\\s", ""),
        # turning numbers into true numbers
        value = as.numeric(value),
        # removing values that are years
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |>
      # removing any remaining duplicate rows
      distinct() |>
      # pivoting to create 'tidy' variables
      pivot_wider(
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |>
      # tidying names
      clean_names() |>
      # removing headings from the table
      select(
        -lederlonn_i_hele_1000,
        -resultatregnskap_i_hele_1000,
        -balanseregnskap_i_hele_1000
      ) |>
      # making it clear what lonn refers to
      rename(
        lederlonn = lonn
      ) |>
      # Creating new variables for plots
      mutate(
        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
      ) |>
      # arranging by year for practical display
      arrange(desc(year))
  } else {
    validate("Please ensure you have entered the correct start and end years")
  }
}

# creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))

    # Testing
    html_test <- read_proff_test(input$html$datapath)

    # Creating data frame

    read_proff_complete(
      html_test,
      as.character(input$start), as.character(input$end)
    )
  })
```

```{r}
html_testing <- function(file, start, end) {
  # Reading file and node
  htmltest <- read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")

  # Checking of node exists
  if (class(htmltest) == "xml_node") {
    htmltable <- htmltest |>
      html_table()
  } else {
    validate("table error")
  }

  # Basic check if the table is correct
  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
    htmlready <- htmltable
  } else {
    validate("regnskap error")
  }

  # Testing if the correct years have been entered
  if (as.character(all_of(start)) %in% names(htmlready) &&
    as.character(all_of(end)) %in% names(htmlready)) {
    # Creating the data frame
    read_proff(file, start, end)
  } else {
    validate("Please ensure you have entered the correct start and end years.")
  }
}

read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |> 
      # removing empty column, using all_of(argument)
      select(
        REGNSKAPSPERIODE:all_of(start)
      ) |>
      # removing white space
      mutate(
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |>
      # removing duplicate table sometimes created by read_html()
      filter(
        !row_number() > 176
      ) |>
      # removing duplicate rows
      filter(
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |>
      # tidying data using pivot_longer() and all_of(argument)
      pivot_longer(
        all_of(end):all_of(start),
        names_to = "year"
      ) |>
      # creating a currency code column and turning current currency codes into
      # NAs
      mutate(
        # changing to real NAs
        value = na_if(value, "-"),
        # turning years into numbers
        year = as.integer(year),
        # creating a currency code column and turning current currency codes into
        # NAs
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |>
      # filling the currency column
      fill(
        valutakode,
        .direction = "updown"
      ) |>
      # removing dates and redundant currency
      filter(
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode"
      ) |>
      mutate(
        # removing white space in numbers
        value = str_replace_all(value, "\\s", ""),
        # turning numbers into true numbers
        value = as.numeric(value),
        # removing values that are years
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |>
      # removing any remaining duplicate rows
      distinct() |>
      # pivoting to create 'tidy' variables
      pivot_wider(
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |>
      # tidying names
      clean_names() |>
      # removing headings from the table
      select(
        -lederlonn_i_hele_1000,
        -resultatregnskap_i_hele_1000,
        -balanseregnskap_i_hele_1000
      ) |>
      # making it clear what lonn refers to
      rename(
        lederlonn = lonn
      ) |>
      # Creating new variables for plots
      mutate(
        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
      ) |>
      # arranging by year for practical display
      arrange(desc(year))
}
```

```{r}
read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column
    fill(
      valutakode,
      .direction = "updown"
    ) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}
```

```{r}
read_proff("Data/Kreftforeningen2002_2021 copy.html", "2002", "2021")

valutatest <- read_html("Data/Kreftforeningen2002_2021 copy.html", encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |> 
   select(
      REGNSKAPSPERIODE:`2002`
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |> 
  pivot_longer(
      `2021`:`2002`,
      names_to = "year"
    ) |> 
   mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |> 
  fill(
      valutakode,
      .direction = "updown"
    ) |>
   distinct() |> 
   pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |> 
  clean_names() |> 
  select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> 
  rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))

valutatest

unique(valutatest$valutakode)

length(unique(valutatest$valutakode))

na.omit(valutatest$valutakode) |> unique()

ifelse(length(unique(valutatest$valutakode)) == 1, TRUE, FALSE)

pull(.data = valutatest, valutakode) |> unique() |> length()

unique(na.omit(pull(.data = valutatest, valutakode)))
```

```{r}
read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column

    ### NEED TO SPLIT HERE

    fill(
      valutakode,
      .direction = "downup"
    ) # |>
    # removing dates and redundant currency
    # filter(
    #   REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
    #     REGNSKAPSPERIODE != "Valutakode"
    # ) |>
    # mutate(
    #   # removing white space in numbers
    #   value = str_replace_all(value, "\\s", ""),
    #   # turning numbers into true numbers
    #   value = as.numeric(value),
    #   # removing values that are years
    #   value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    # ) |>
    # # removing any remaining duplicate rows
    # distinct() |>
    # # pivoting to create 'tidy' variables
    # pivot_wider(
    #   names_from = REGNSKAPSPERIODE, values_from = value
    # ) |>
    # # tidying names
    # clean_names() |>
    # # removing headings from the table
    # select(
    #   -lederlonn_i_hele_1000,
    #   -resultatregnskap_i_hele_1000,
    #   -balanseregnskap_i_hele_1000
    # ) |>
    # # making it clear what lonn refers to
    # rename(
    #   lederlonn = lonn
    # ) |>
    # # Creating new variables for plots
    # mutate(
    #   annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
    #   balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    # ) |>
    # # arranging by year for practical display
    # arrange(desc(year))
}

test <- read_proff("Data/Kreftforeningen2002_2021 copy.html", "2002", "2021")

unique(test$valutakode)

test
```