### 2022-08-01 Report/KPIs and Visualisations for report.Rmd

---
title: "KPIs and Visualisations for report"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-08-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Key Performance Indicators

On the basis of such data as we could gather we developed five potential KPIs:

- **Fundraising ROI**:
  - A basic measure of the return on investment of direct fundraising expenses:

  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$

This KPI is not a true measure of fundraising ROI, as it only includes (as we understand OPX's data) the direct fundraising expenses, and does not include major indirect expenses such as salary costs.

***

- **Liquid assets/obligations**:
  - A measure of the relationship between the organisation's liquid assets and its obligations:

  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$

This KPI was suggested by OPX.

***

- **Private donations growth**:
  - A calculation of the percentage change in private contributions based on the two most recent years:

  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$

***

- **Public grants growth**:
  - A calculation of the percentage change in public grants received based on the two most recent years:

$$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$

***

- **Private/public proportion**:
  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:

$$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$

This KPI may say something about the NGO's sources of funding. There are prima facie no specific desirable funding mix, but this measure may be of interest to some donors.

***

### Visualisations

The dashboard includes five plots, three based on data from Proff.no and two based on data from the Excel sheet.

-   **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
-   **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
-   **Revenues**: Based on the category *Revenue* in the Excel sheet.
-   **Expenses**: Based on the category *Expenses* in the Excel sheet.
-   **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
    -   This appears to increasingly be treated as a necessary measure of transparency and performance, but the values should normally be evaluated in the context of the size of the NGO and the sector(s) within which it operates.

The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.