### 2022-08-02 Dashboard/Dashboard testing material/zero testing.Rmd

---
title: "Zero testing"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/ISSSV1337/ISSSV1337-Case/Dashboard")
```

```{r}
x <- c(5, NA, 0, 0, 10, -15)

sum(x == 0, na.rm = TRUE)

warning_message <- "ALl values are zero. This is unusual and may indicate a data problem."

?pull

setwd("~/ISSSV1337/ISSSV1337-Case/Dashboard")

exceltest <- read_excel("Data/Kreftforeningen.xlsx") |> clean_names()

exceltest

revenue_zero <- exceltest |>
  filter(category == "Revenue") |>
  summarise(
    zero = sum(value != 0, na.rm = TRUE)
  ) |> 
  pull()

revenue_zero

excel_zero_check <- function(type) {
  exceltest |>
    filter(
      category == as.character(all_of(type))
    ) |>
    summarise(
      zero = sum(value != 0, na.rm = TRUE)
    ) |>
    pull()
}

excel_zero_check("Revenue")
```