### 2022-08-03 Presentasjon/presentasjon.Rmd
Nye stikkord

---
title: "Presentasjon OPX Fiffig tittel"
author: "Gruppe 5"
date: "4.8.2022"
output: 
  ioslides_presentation: 
    css: stylesss.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

## <br> OPX and their goal
- Develop solutions for socially responsible investment strategies
- Increase cash flow to charitable purposes <br>

```{r, out.height="128px", out.width="568px", fig.align='center'}
knitr::include_graphics("Images/logo_opx.png")
```

## <br> How to assess financial performance?
- Data collection 
- Key Performance Indicators <br>

```{r, out.height="268px", out.width="350px", fig.align='center'}
knitr::include_graphics("Images/kpi.png")

```

## {data-background="Images/brainstorming2.png" data-background-size=cover}

## <br> Phase 1 – how to harvest and analyze the data?
- OPX wanted us to harvest data --> how? <br> 
- A webcrawler? <br> 
- No: Would not work for every organization <br> 
- A code that could read PDF? <br> 
- Perhaps: Financial statements as text tables and as images

## <br> Phase 2 – Exploring three options
- 1. Reading PDFs <br>
Issue: Not always published as text
- 2. Webscraping from Proff.no <br>
Issue: Proff.no does not allow webscraping
- 3. Webscraping the Brønnøysund registry <br>
Issue: No API access, and access costs ~ kr 500K

## <br> Workarounds... 
- Proff.no: Download HTML?
- Brønnøysund: Request temporary access

## <br> ... and why they did not work
- Different accounting items used
- Incomplete financial reporting from NGOs

- **Which led us to Phase 3 - scrapping the scraping**

----
```{r, out.height="360", out.width="612", fig.align='center'}
knitr::include_graphics("Images/fcuk.png")

```

## <br> Phase 3 – Scrapping the scraping
- OPX wanted KPIs <br>
- Found useful KPIs <br>
- Created an interactive dashboard

## <br> Data collection & data description
- Used data from OPX and Proff.no <br>
- OPX: From their webpage. <br>
- Used to create KPIs and check our codes <br>
- Proff.no: Downloaded HTML documents. <br>
- Used to create visualizations.

# Findings

# Limitations and steps forward

## <iframe width="1280" height="720" src="Images/henrik.mp4" align="middle" frameborder="0" allowfullscreen></iframe>

## {data-background="Images/tenke.png" data-background-size=cover}

## <br> Our experience
- Unsolvable case due to non-existent data
- Hard to create credible KPIs due to limited knowledge of economics and the NGO sector
- **Insight into real-life work challenges - the clash between what is desired and what is possible**